\documentclass[12pt]{article}
\input{../preamble}

% ---------------------------------------------
\title{Crowded Primaries and Weakened Nominees: Behavioral Distortion in a Two-Stage Electoral Contest}
\author{Katherine Parslow}
\date{Last Updated: \today}

\begin{document}
\maketitle

\section*{Model \& Notation}

\subsection*{Setup}

$N \ge 2$ candidates compete for a party's nomination.
Each candidate $i$ has valence $v_i \in \R$, drawn i.i.d.\ from distribution $F$
with strictly positive density $f$ on $\R$.
The valence profile $\vprof = (v_1,\dots,v_N)$ is drawn by nature at the start of the game
and is observed by all candidates.

Candidates simultaneously choose an action $a_i \in \{0,1\}$.
Taking action ($a_i = 1$) boosts a candidate's relative standing in the primary
but harms their viability in the subsequent general election.
Let $\aprof = (a_1,\dots,a_N)$ denote the action profile.

\subsection*{Primary Election}

Candidates are ranked according to their \emph{primary index}
\begin{equation}
    q_i^P = v_i + a_i + \varepsilon_i,
    \label{eq:primary-index}
\end{equation}
where $\varepsilon_i \sim \text{EV1}$ are i.i.d.\ candidate-level shocks.
The nominee, denoted $w$, is the candidate with the highest primary index.

Define the \emph{opponent-strength aggregator} for candidate $i$ as
\[
    \Sof{\vminus}{\aminus} = \sum_{j \neq i} \ex{v_j + a_j}.
\]
Integrating over the EV1 shocks, the probability that candidate $i$ wins the nomination is
\begin{equation}
    \Pr(w = i \mid \vprof, \aprof)
        = \pnomrf{a_i}{v_i}{\Sof{\vminus}{\aminus}}.
    \label{eq:primary-prob}
\end{equation}
We write this more compactly as $\pnom{a_i}{v_i}{\Sof{\vminus}{\aminus}}$,
with the special cases $\pOne{v_i}{\Sof{\vminus}{\aminus}}$ when $a_i = 1$
and $\pZero{v_i}{\Sof{\vminus}{\aminus}}$ when $a_i = 0$.

\subsection*{General Election}

The nominee $w$ advances to the general election.
Their \emph{general-election index} is
\begin{equation}
    q_w^{\GE} = v_w - \delta\, a_w + \eta_w,
    \label{eq:ge-index}
\end{equation}
where $\eta_w \sim N(0,1)$ is an idiosyncratic shock independent of the primary shocks,
and $\delta > 0$ captures the cost of primary electioneering.
Normalizing the out-party opponent's index to $0$, the probability that nominee $w$ wins the general election is
\begin{equation}
    \Pr(\GE\text{ win} \mid v_w, a_w) = \Pge{v_w}{\delta}{a_w}.
    \label{eq:ge-prob}
\end{equation}

\subsection*{Payoffs}

Candidates are office-motivated and receive a payoff of $1$ from holding office and $0$ otherwise.
Because holding office requires winning both elections, candidate $i$'s expected utility is
\begin{equation}
    u_i(\vprof, \aprof)
        = \pnom{a_i}{v_i}{\Sof{\vminus}{\aminus}} \cdot \Pge{v_i}{\delta}{a_i}.
    \label{eq:utility}
\end{equation}

\subsection*{Timing}

The timing of the game is as follows:
\begin{enumerate}
    \item Nature draws the valence profile $\vprof = (v_1,\dots,v_N)$;
    \item Candidates observe $\vprof$ and simultaneously choose $a_i \in \{0,1\}$ for all $i$;
    \item The primary election is held and the party nominee $w$ is selected according to \eqref{eq:primary-prob};
    \item The general election is held and payoffs are determined according to \eqref{eq:ge-prob}.
\end{enumerate}

\subsection*{Notation Summary}

\begin{center}
\begin{tabular}{cl}
    \hline
    \textbf{Symbol} & \textbf{Description} \\
    \hline
    $N$ & Number of candidates \\
    $v_i$ & Valence of candidate $i$; drawn i.i.d.\ from $F$ \\
    $\vprof = (v_1,\dots,v_N)$ & Valence profile \\
    $a_i \in \{0,1\}$ & Candidate $i$'s primary-electioneering action \\
    $\aprof = (a_1,\dots,a_N)$ & Action profile \\
    $\delta > 0$ & Cost of taking the primary action in the general election \\
    $\Sof{\vminus}{\aminus}$ & Opponent-strength aggregator $\sum_{j\neq i}\exp(v_j+a_j)$ \\
    $\pnom{a}{v}{S}$ & Nomination probability for action $a$, valence $v$, aggregator $S$ \\
    $\Phi(\cdot)$ & Standard normal CDF; used for GE win probability \\
    \hline
\end{tabular}
\end{center}

\end{document}